---
title: "Regression Model Course Project"
author: "Ersi Ni"
date: "Jan 20. 2015"
output: rmarkdown::tufte_handout
---

# Introduction

This is the report to my hyperthetical work at *Motor Trend*, looking at the data of a collection of cars. This report delivers the analysis over whether Automatic Transmission or Manual Transmission is better for Miles Per Gallon.

# Anlaysis

```{r, echo=FALSE}
library(datasets)
data(mtcars)
```

A short description of what each variable means in the `mtcars` data set would act as a foreword for this report before we go into detailed analysis:

  * mpg -> our outcome, miles per gallon
  * cyl = numbers of cylinders
  * disp = displacement
  * hp = horse power
  * wt = weight
  * am = our predictor in question, transmission mode (0 for automatic, 1 for manual)
  * gear = number of forward gear
  * carb = number of carburetors


First we look at the linear model for the outcome `mpg` from all predictor variables inside the data set `mtcars`, specifically we look at the coefficients of the linear model:

```{r, echo = F, results='asis'}
library(xtable)
options(xtable.comment = FALSE)
options(xtable.booktabs = TRUE)
xtable(summary(lm(mpg~.,data=mtcars))$coefficients, caption = "P value of linear regression")
```


We noticed that the p value of the coefficients will give us clue as to how much the variable is a factor to the outcome `mpg`. Because the p-value tests the null hypothesis that the coefficient is equal to zero. A low (<0.05) p-value indicates that we can reject this null hypothesis, thus a meaningful variable to our model. Surprisingly our questioned `am` variable is not the lowest of them, but `am` is lower than most others. That means we need to dig deeper.

To support the initial assumption that `am` is indeed a factor of `mpg` response,
we can view the plot of transmission mode over `mpg`.

```{r, echo=FALSE, fig.cap = "Transmission Mode on Miles per Gallon"}
library(ggplot2)
g <- ggplot(mtcars, aes(factor(am),mpg, fill=factor(am)))
g = g +
  geom_boxplot() +
  geom_jitter(position = position_jitter(width = .1, height = 0)) +
  scale_color_discrete(name = "Transmission Mode") +
  scale_fill_discrete(name = "Transmission Mode", breaks = c("0", "1"), labels = c("Automatic", "Manual")) +
  scale_x_discrete(breaks = c("0", "1"), labels = c("Automatic", "Manual")) +
  xlab("")
g

```


From the figure we can see the obvious difference in mean between 2 modes impacting on `mpg`.
We make an hyperthesis test that Ho is `mpg` of manual is less than mpg of automatic, while Ha
is `mpg` of manual is greater than mpg of automatic.

```{r, echo=F}
m1 <- subset(mtcars,mtcars$am==0)
m2 <- subset(mtcars,mtcars$am==1)
t.test(m1,m2,alternative="greater",paired=F)
```

*Now we can say manual transmission has better mpg performance.*

# Quantify the MPG difference

Going back to the first linear model we used at the begining, we know that that summary gives us a clue which variables might be thrown away because they don't appear to be a confounding factor. Step-wise model can deliver that.

```{r}
step_lm <- step(lm(data=mtcars, mpg ~ .), trace = 0)
summary(step_lm)
```

The step model gives us 3 variables `wt`, `qsec` and `am` that can yield a better R-Sqaure. 

Our goal is to find a model with least regressor and highest R-sqaure, the strategy here is comparing several models. To yield a significant difference, we can pick the original all variable model, a model with variables that are highly correlated, a model of `wt` alone and the step-wise model.

The highest correlated variables are 
```{r, echo=FALSE}
head(names(sort(abs(round(cor(mtcars)[1,],2)),decreasing=T)))
```
So we have 

1. `wt+qsec+am` model `r fit1 <- lm(data=mtcars, mpg~wt+qsec+am)`
4. all variable model `r fit2 <- lm(data=mtcars, mpg~.)`
3. `wt` model `r fit3 <- lm(data=mtcars, mpg~wt)`
2. high correlated `wt+cyl+disp+hp` model `r fit4 <- lm(data=mtcars, mpg~wt+cyl+disp+hp)`

Now we can compare them by using `anova`

```{r}
anova(fit1, fit2, fit3, fit4)
```

```{r, fig.margin=T}
r1 <- resid(fit1)
plot(r1, main="model wt+qsec+am")
abline(h=0,lwd=2)
```

# Headings

This style provides a- and b-heads (that is, `#` and `##`), demonstrated above.
An error is emitted if you try to use `###` and smaller headings.

\newthought{In his later books}[^books_be], Tufte starts each section with a bit of vertical space, a non-indented paragraph, and sets the first few words of the sentence in small caps. To accomplish this using this style, use the `\newthought` command as demonstrated at the beginning of this paragraph.

# Figures

## Margin Figures

Images and graphics play an integral role in Tufte's work. To place figures or tables in the margin you can use the `fig.margin` knitr chunk option. For example:

```{r, fig.margin = TRUE, fig.cap = "Sepal length vs. petal length, colored by species"}
library(ggplot2)
qplot(Sepal.Length, Petal.Length, data = iris, color = Species)
```

Note the use of the `fig.cap` chunk option to provide a figure caption. You can adjust the proportions of figures using the `fig.width` and `fig.height` chunk options. These are specified in inches, and will be automatically scaled down to fit within the handout margin.

## Equations

You can also include \LaTeX\ equations in the margin by explicitly invoking the `marginfigure` environment.

\begin{marginfigure}
$$\frac{d}{dx}\left( \int_{0}^{x} f(u)\,du\right)=f(x).$$
\caption{An equation}
\end{marginfigure}

Note the use of the `\caption` command to add additional text below the equation.

## Full Width Figures

You can arrange for figures to span across the entire page by using the `fig.fullwidth` chunk option. 

```{r, fig.width = 10, fig.height = 2, fig.fullwidth = TRUE, fig.cap = "Full width figure"}
qplot(wt, mpg, data=mtcars, colour=factor(cyl))
```

Note the use of the `fig.width` and `fig.height` chunk options to establish the proportions of the figure. Full width figures look much better if their height is minimized.

## Main Column Figures

Besides margin and full width figures, you can of course also include figures constrained to the main column.

```{r, fig.cap = "Another figure"}
qplot(factor(cyl), mpg, data = mtcars, geom = "boxplot")
```

# Sidenotes

One of the most prominent and distinctive features of this style is the extensive use of sidenotes. There is a wide margin to provide ample room for sidenotes and small figures. Any use of a footnote will automatically be converted to a sidenote. ^[This is a sidenote that was entered using a footnote.] 

If you'd like to place ancillary information in the margin without the sidenote mark (the superscript number), you can use the `\marginnote` command. \marginnote{This is a margin note.  Notice that there isn't a number preceding the note.}

Note also that the two footnote references (`tufte_latex` and `books_be`, both defined below) were also included in the margin on the first page of this document.

# Tables

You can use the **xtable** package to format \LaTeX\ tables that integrate well with the rest of the Tufte handout style. Note that it's important to set the `xtable.comment` and `xtable.booktabs` options as shown below to ensure the table is formatted correctly for inclusion in the document.

```{r, results='asis'}
library(xtable)
options(xtable.comment = FALSE)
options(xtable.booktabs = TRUE)
xtable(head(mtcars[,1:6]), caption = "First rows of mtcars")
```


```{r, echo=FALSE,eval=FALSE}
dat <- mtcars[,c("mpg","wt","qsec","am")]
fit <- lm(mpg ~ . - 1, data = dat)
summary(fit)$coefficients

pairs(dat, panel= panel.smooth, main="mtcars data")
par(mfrow = c(2,2))
plot(fit)
sumCoef <- summary(fit)$coefficients
interval <- sumCoef["am",1]+c(-1,1)*qt(0.975,df=fit$df)*sumCoef["am",2]
interval

```


[^tufte_latex]: https://code.google.com/p/tufte-latex/
[^books_be]: http://www.edwardtufte.com/tufte/books_be










